---
title: README
output: github_document
---

**Comparing uncertainty associated with 1-, 2-, and 3D aerial photogrammetry-based body condition measurements of baleen whales**   
  *Data and model code*



*Steps:*   
1. Run uncertainty model on total length and width measurements to produce a posterior predictive distribution for each measurement of each whale.     
2. Calculate the different body condition measurements (1D, 2D,and 3D) for each individual using the uncertainty model outputs (the posterior predictive distributions for each measurement).        
3. Analyze and compare the different body condition measurements.      

&nbsp;

# 1. Running the uncertainty model

Uncertainty model from: https://github.com/KCBierlich/Uncertainty_Model    

The code is designed to be run using the ```drake``` package, by running the 
code in the script [make.R](make.R) either via the command line 
(i.e., ```R CMD BATCH make.R```), or from within an interactive ```R``` session.
The outline of all project components (i.e., the ```drake``` plan) is 
available in the [R/subplans](R/subplans) directory.

The ```drake``` [Github page](https://github.com/ropensci/drake) page has a 
good overview of what drake does, and includes some code snippets.  After 
running the [make.R](make.R) script, the ```drake::loadd``` function is the 
most helpful function for loading output.

The ```drake``` [documentation](https://books.ropensci.org/drake/index.html) 
itself is also a good resource for learning how to use the ```drake``` package.  
Starting with the 
[Walkthrough](https://books.ropensci.org/drake/walkthrough.html#set-the-stage.)
is a decent place to just jump in to some details.


## Data formats

The Error measurement model is implemented in the 
[model_joint.R](R/nimble/model_joint.R) script.  The model implementation is 
abstract, to allow multiple observations of any object.  The 
[data_plan.R](R/subplans/data_plan.R) script creates the necessary, raw data 
structures for the model.  The [flatten_data()](R/nimble/flatten_data.R)
function munges the data structures for use with the model's implementation
in ```nimble```.


### Training object information

Each training object must be documented in the same format as the 
```training_obj``` object (below).  The first two columns specify the 
training object, and the third column records its known length.  The 
[flatten_data()](R/nimble/flatten_data.R)
function will use the information in ```training_obj``` to 
associate known lengths with image and pixel measurements of the training 
objects.


```{r, echo = FALSE}
library(drake)
loadd(training_obj)
knitr::kable(training_obj, row.names = FALSE)
```


### Image data

Each image must be documented in the same format as the ```Mns_images``` object
(preview below).  The first column ```Image``` provides a unique identifier for 
the image and will be used to create an altitude variable, which will be 
estimated from the altitude sensor readings provided by ```AltitudeBarometer``` 
and ```AltitudeLaser``` in conjunction with training data.  The remaining 
columns in ```Mns_images``` report the focal length (mm), image width (pixels), 
and camera sensor width (mm) associated with the image, which will be used to 
estimate the image's ground sampling distance (GSD) per pixel.

```{r, echo = FALSE}
library(drake)
loadd(Mns_images)
knitr::kable(Mns_images[1:5,])
```

### Pixel count "measurement" data

All pixel measurements must also be documented in the same format as the 
```Mns_pixels``` object (preview below).  The first two columns specify the 
measurement.  One measurement variable, which will be estimated, will be 
associated with each unique combination of the first two columns. The 
```Image``` column links the measurement to an image and it's estimated GSD.
Lastly, the ```PixelCount``` column records the pixel-length of the object as 
it appears in the image.  

```{r, echo = FALSE}
library(drake)
loadd(Mns_pixels)
knitr::kable(Mns_pixels[1:5,])
```

### Many-to-Many relationships

Storing Image and Pixel count data in separate structures allows multiple 
measurements to be estimated from a single image.  For example, total length and 
maximum width can be measured from the same image through a pixel count table 
like the following:

```{r, echo = FALSE}
knitr::kable(
  data.frame(Subject = 'Animal A', Measurement = c('Total length', 'Max. width'), 
           Image = 'Image 1', PixelCount = c(1e3, 1e2))
)

```

Both ```PixelCount``` entries above relate to a different object being measured, 
but use the same estimated altitude and implied GSD for ```Image 1``` when
estimating lengths from pixel counts.


Similarly, the Image and Pixel count structures also allow one object to be 
estimated from multiple images.  For example, total length can be estimated 
from multiple observations through a pixel count table like the following:

```{r, echo = FALSE}
knitr::kable(
  data.frame(Subject = 'Animal A', Measurement = 'Total length', 
           Image = paste('Image', 1:5, sep = ' '), 
           PixelCount = c(1e3, rnorm(n = 4, mean = 1e3, sd = 20)))
)
```


## Data
Data located in data folder  

_calibration_object-measurements.csv_ - training/calibration data for the LemHex-44 and FreeFly Alta 6.     
Data from: Bierlich, K. C., Schick, R. S., Hewitt, J., Dale, J., Goldbogen, J. A., Friedlaender, A. S., & Johnston, D. W. (2020). Data and scripts from: A Bayesian approach for predicting photogrammetric uncertainty in morphometric measurements derived from UAS. Duke Research Data Repository. V2 https://doi.org/10.7924/r4sj1jj6s  

  * CO.ID = calibration object ID  
  * CO.Length = the true length of the CO.ID  
  * Image = image used for measuring calibration object  
  * Lab = the lab that collected the data  
  * Cruise = research expedicition ID  
  * Date = date that imagery was collected of calbiration object  
  * Flight = the flight number/name  
  * Pilot = pilot during data collection  
  * VO = visual observer  
  * Aircraft = the UAS aircraft used to collect imagery of calibration object to measure  
  * Focal_length = focal length of camera  
  * Iw = image width in pixels  
  * Sw = sensor width in mm  
  * pix.dim = pixel dimensions; Sw/Iw
  * Baro_raw = the raw relative altitude recorded by the barometer  
  * Launch.Ht = the launch height of the drone, to be added to the BarAlt to get the absolute barometric altitude above the surface of the water  
  * Baro+Ht = the baro_raw + Launch.Ht to get the absolute barometer altitude  
  * Laser_Alt = the altitude recorded by the laser altimeter. Blanks spaces/NAs indicate no/false reading
  * Altitude = altitude used in measurement, either Laser_Alt or Baro + Ht  
  * Altimeter = which altimeter was used for altitude in measurement; either barometer or laser  
  * Lpix = the length in pixels of the known sized object (calibration object)    
  * object_position = indicates if calibration object is in center of corner of image frame  
  * Analyst = analyst that performed the measurement  
  
  
_measurement_inputs.csv_ - total length and body width measurements for blue, humpback, and Antarctic minke whales.  Measurements performed using [MorphoMetriX](https://github.com/wingtorres/morphometrix)   

  * AID = unique animal ID for individual whale    
  * Species = the species of whale   
  * Image = the image ID used for measuring the whale  
  * TL = total length of the whale in meters, measure rostrum to fluke notch  
  * TL.05.00..Width - TL.95.00..Width = width measurements in % increments of TL    
  * Reproductive_Class = actually pretty limited here, just "calf" or "adult"   
  * HTRange = Head to Tail Range used to calculate 2D and 3D body condition measurements  
  * SHrange = Short Range, only includes the three widths with the largest standard deviation for the population. Not included in the analysis  
  * SW = the Single Width (1D) measurement, width with the largest standard deviation for population  
  * BaroAlt = the raw relative altitude recorded by the barometer  
  * Launch_Ht = the launch height of the drone, to be added to the BarAlt to get the absolute barometric altitude  
  * LaserAlt = the altitude recorded by the laser altimeter. Blanks spaces/NAs indicate no/false reading  
  * Altitude = altitude used in measurement, either LaserAlt or BaroAlt + Launch_Ht  
  * Focal_Length = focal length (mm) of the camera used  
  * Iw = image width, in pixels  
  * Sw = sensor width, in mm  
  * alt_diff = percent difference between BaroAlt + Launch_Ht and LaserAlt  
 
 
  
## Steps
run 'make.R'  

This will generate an "output" folder that contains an 'mcmc' folder and a 'reports' folder.   
 
  * The 'mcmc' folder contains all the .rds outputs from the model.  
      + "length_samples_mns.rds" contains the posterior predictive measurement distributions for TL and widths and is used to calculate  body condition in the "calculating_body_condition" folder. 
  
  * The 'reports' folder contains .html files to evaluate the results from the model  
      + See "posterior_diagnostics_mns.html" for a summary of model output for each measurement

After model is finished running, proceed to calculate_body condition      
      

# 2. Calculating body condition  

In the "calculate_body_condition" folder, run "Calculate_body_condition.Rmd".    
  
This uses [CollatriX](https://github.com/cbirdferrer/collatrix#installation) to calculate body condition measurements.

_Creates:_    
  * whales_merged.csv -- cleaned dataframe to input into CollatriX  
  * collated_allMC.rds -- each animal's dataframe containing the posterior distribution (the mcmc iterations) for each measurement  
  * collated_MC_summarystats.csv -- CollatriX summary sheet with the mean and 95% HPD intervals for the posterior predictive distribution of each body condition metric for each individual. Used in body condition analysis. 



# 3. body condition analysis

Run "Body_condition_analysis.Rmd"

This will run the analysis comparing the different body condition metrics and re-create figures from manuscript.




## Contact
[KC Bierlich](https://github.com/KCBierlich), kevin.bierlich@oregonstate.edu  

Core contributors:  
[Clara N. Bird](https://github.com/cbirdferrer)  
[Dr. Josh Hewitt](https://github.com/jmhewitt)  

