---
title: "body_condition_analysis"
author: "KCB"
date: "7/29/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Comparing 1D, 2D, and 3D body condition measurements {.tabset}
This R markdown re-runs the analysis and creates the figures in the manuscript.   

```{r packages, include = FALSE}
library(ggplot2)
library(conflicted)
library(dotenv)
library(drake)
library(dplyr)
library(tidyr)
library(ggpubr)
library(coda)
library(ggridges) # ridgeplots
library(ggdist) # Using TidyBayes
library(corrplot) # Correlation Matrix
library(conflicted) 
conflict_prefer("lag", "dplyr") 
conflict_prefer("select", "dplyr")
conflict_prefer("mutate", "dplyr") 
conflict_prefer("summarise", "dplyr") 
conflict_prefer("summarize", "dplyr") 
```


## Import / prep data 
```{r}
# load each animal's dataframe, posterior distribution (the mcmc iterations)
indv_df <- readRDS(file.path("..","calculate_body_condition", "outputs", "collated_all_iter.rds"))


# CollatriX summary sheet with the mean and 95% HPD intervals for the posterior predictive distribution of each body condition metric for each individual
df.summary <- read.csv(file.path("..","calculate_body_condition", "outputs", "collated_MC_summarystats.csv"))

# Import model input measurement data
indv_meta <- read.csv(file.path("..","data","measurement_inputs.csv"))
  # keep unique metadata per AID
  indv_meta <- indv_meta %>% group_by(AID) %>% dplyr::filter(row_number(AID) == 1)
  
# Merge CollatriX summary df w/ the metadata associated w/ each image and AID from the input measurements.
df.summary_meta <- df.summary %>% left_join(indv_meta, by = "AID") %>% 
  select(AID, Species, Image, TL, TL.05.00..Width, TL.10.00..Width, TL.15.00..Width, TL.20.00..Width, 
  TL.25.00..Width, TL.30.00..Width, TL.35.00..Width, TL.40.00..Width, TL.45.00..Width, TL.50.00..Width,
  TL.55.00..Width, TL.60.00..Width, TL.65.00..Width, TL.70.00..Width, TL.75.00..Width, TL.80.00..Width,
  TL.85.00..Width, TL.90.00..Width, TL.95.00..Width, HTRange, SW, BaroAlt, Launch_Ht, LaserAlt, Altitude, 
  Focal_Length, Iw, Sw, date_source, Reproductive_Class, TL.mean, TL.lower, TL.upper, SWstd.mean, SWstd.lower, SWstd.upper, BAIpar.HT.mean,
  BAIpar.HT.lower, BAIpar.HT.upper, SA.HT.mean, SA.HT.lower, SA.HT.upper, BV.HT.mean, BV.HT.lower, BV.HT.upper)
# Note that 'TL' is the original/expected/empirical/pre-model length, where 'TL.mean' is the mean of the predicted posterior distribution from the model. Same for the widths.

# add altimeter as a column
bc_df <- rbind(df.summary_meta %>% dplyr::filter(!is.na(BaroAlt) & !is.na(LaserAlt)) %>% mutate(altimeter = "barometer & laser"),
              df.summary_meta %>% dplyr::filter(is.na(BaroAlt)) %>% mutate(altimeter = "only laser"),
              df.summary_meta %>% dplyr::filter(is.na(LaserAlt)) %>% mutate(altimeter = "only barometer"))

# Quick summary
table(bc_df$Species)
table(bc_df$altimeter)
```


#### Pull SWs for each group in analysis
CollatriX calculated SWstd, need to pull the mean of the SW posterior predictive measurements
```{r}

# Calculating SW for each species

# combine the two full df lists
mcmc_list <- c(indv_df)

# save list of AIDs
ID_list <- names(mcmc_list)  # extract AIDs
full_df <- mcmc_list        # save individual's dataframes to be used in the loop for maturity
           
# make an empty dataframe to fill with the summary stats
new_SW.df <- data.frame()

# loop through list of AIDs (loop through each individual's data frame) and calculate the proportion of TLs > 11.2. 
# result has a column of 'perAdult' for percent likelihood of being an adult.
for (x in 1:length(ID_list)){
  xx <- ID_list[x] #pull ID name from list
  AID <- xx #save ID as variable
  TL <- mean(full_df[[xx]]$"TL")

  # for blues
  sw55 <- mean(full_df[[xx]]$"TL.55.00..Width")
  sw55_lwr <- HPDinterval(mcmc(full_df[[xx]]$"TL.55.00..Width"))[1]
  sw55_upr <- HPDinterval(mcmc(full_df[[xx]]$"TL.55.00..Width"))[2]
  sw55_std <- mean(full_df[[xx]]$"TL.55.00..Width"/full_df[[xx]]$"TL")
  sw55_std_lwr <- HPDinterval(mcmc(full_df[[xx]]$"TL.55.00..Width"/full_df[[xx]]$"TL"))[1]
  sw55_std_upr <- HPDinterval(mcmc(full_df[[xx]]$"TL.55.00..Width"/full_df[[xx]]$"TL"))[2]
  
  # for minkes
  sw40 <- mean(full_df[[xx]]$"TL.40.00..Width")
  sw40_lwr <- HPDinterval(mcmc(full_df[[xx]]$"TL.40.00..Width"))[1]
  sw40_upr <- HPDinterval(mcmc(full_df[[xx]]$"TL.40.00..Width"))[2]
  sw40_std <- mean(full_df[[xx]]$"TL.40.00..Width"/full_df[[xx]]$"TL")
  sw40_std_lwr <- HPDinterval(mcmc(full_df[[xx]]$"TL.40.00..Width"/full_df[[xx]]$"TL"))[1]
  sw40_std_upr <- HPDinterval(mcmc(full_df[[xx]]$"TL.40.00..Width"/full_df[[xx]]$"TL"))[2]
  
  
  # for humpback calves
  sw60 <- mean(full_df[[xx]]$"TL.60.00..Width")
  sw60_lwr <- HPDinterval(mcmc(full_df[[xx]]$"TL.60.00..Width"))[1]
  sw60_upr <- HPDinterval(mcmc(full_df[[xx]]$"TL.60.00..Width"))[2]
  sw60_std <- mean(full_df[[xx]]$"TL.60.00..Width"/full_df[[xx]]$"TL")
  sw60_std_lwr <- HPDinterval(mcmc(full_df[[xx]]$"TL.60.00..Width"/full_df[[xx]]$"TL"))[1]
  sw60_std_upr <- HPDinterval(mcmc(full_df[[xx]]$"TL.60.00..Width"/full_df[[xx]]$"TL"))[2]
  
  # for humpback adults
  sw50 <- mean(full_df[[xx]]$"TL.50.00..Width")
  sw50_lwr <- HPDinterval(mcmc(full_df[[xx]]$"TL.50.00..Width"))[1]
  sw50_upr <- HPDinterval(mcmc(full_df[[xx]]$"TL.50.00..Width"))[2]
  sw50_std <- mean(full_df[[xx]]$"TL.50.00..Width"/full_df[[xx]]$"TL")
  sw50_std_lwr <- HPDinterval(mcmc(full_df[[xx]]$"TL.50.00..Width"/full_df[[xx]]$"TL"))[1]
  sw50_std_upr <- HPDinterval(mcmc(full_df[[xx]]$"TL.50.00..Width"/full_df[[xx]]$"TL"))[2]
  
  
  # standard deviations of uncertainty
  #{!} standard error or variance or deviation
  TL.sd <- sd(full_df[[xx]]$TL)
  BAI.sd <- sd(full_df[[xx]]$BAIpar_HT)
  SA.sd <- sd(full_df[[xx]]$SA_HT)
  BV.sd <- sd(full_df[[xx]]$BV_HT)
  sw55.sd <- sd(full_df[[xx]]$"TL.55.00..Width")
  sw40.sd <- sd(full_df[[xx]]$"TL.40.00..Width")
  sw60.sd <- sd(full_df[[xx]]$"TL.60.00..Width")
  sw50.sd <- sd(full_df[[xx]]$"TL.50.00..Width")
  sw55std.sd <- sd(full_df[[xx]]$"TL.55.00..Width"/full_df[[xx]]$"TL")
  sw40std.sd <- sd(full_df[[xx]]$"TL.40.00..Width"/full_df[[xx]]$"TL")
  sw60std.sd <- sd(full_df[[xx]]$"TL.60.00..Width"/full_df[[xx]]$"TL")
  sw50std.sd <- sd(full_df[[xx]]$"TL.50.00..Width"/full_df[[xx]]$"TL")
  
  # merge
  tempdf <- data.frame(AID, TL,
                       sw55, sw55_lwr, sw55_upr, sw55_std, sw55_std_lwr, sw55_std_upr,
                       sw40, sw40_lwr, sw40_upr, sw40_std, sw40_std_lwr, sw40_std_upr,
                       sw60, sw60_lwr, sw60_upr, sw60_std, sw60_std_lwr, sw60_std_upr,
                       sw50, sw50_lwr, sw50_upr, sw50_std, sw50_std_lwr, sw50_std_upr,
                       TL.sd, BAI.sd, SA.sd, BV.sd, sw55.sd, sw40.sd, sw60.sd, sw50.sd,
                       sw55std.sd, sw40std.sd, sw60std.sd, sw50std.sd)
  new_SW.df <- rbind(new_SW.df,tempdf) #add the row of summary stats for this individual to the main dataframe
}

new_SW.df
```

Now merge new SW w/ dataframe, and clean up
```{r}

bc_dat <- bc_df %>% left_join(new_SW.df %>% select(!c(TL)), by = "AID")
bc_dat <- bc_dat %>% select(!"TL.95.00..Width")  # don't need this width


# create a group to distinguish bluew, AMWs, humpbacks, and humpback calves
bc_dat$group <- bc_dat$Species
bc_dat$group <- ordered(bc_dat$group, levels = c("AMW", "Blue", "Humpback", "Humpback calf"))
bc_dat$group[bc_dat$Reproductive_Class == "calf"]  <- 'Humpback calf'


# Consistently color by species 
specColors <- setNames(c("aquamarine3", "steelblue", "coral3", "coral"), names(bc_dat$group))


# now need to add the standardized SW for each species 
std_SW <- rbind(bc_dat %>% 
                  dplyr::filter(Species == "Humpback" & Reproductive_Class != "calf") %>% 
                  mutate(SW = sw50, SW_lwr = sw50_lwr, SW_upr = sw50_upr, 
                         SW_std = sw50_std,SW_std_lwr = sw50_std_lwr, 
                         SW_std_upr = sw50_std_upr, SW_sd = sw50.sd, SW_std_sd = sw50std.sd) %>% 
                  select(AID, SW, SW_lwr, SW_upr, SW_sd, SW_std, SW_std_lwr, SW_std_upr, SW_std_sd),
                
                
                bc_dat %>% 
                  dplyr::filter(group == "Humpback calf") %>% 
                  mutate(SW = sw60, SW_lwr = sw60_lwr, SW_upr = sw60_upr, 
                         SW_std = sw60_std,SW_std_lwr = sw60_std_lwr, 
                         SW_std_upr = sw60_std_upr, SW_sd = sw60.sd, SW_std_sd = sw60std.sd) %>% 
                  select(AID, SW, SW_lwr, SW_upr, SW_sd, SW_std, SW_std_lwr, SW_std_upr, SW_std_sd),
      
                
                bc_dat %>% 
                  dplyr::filter(Species == "Blue") %>% 
                  mutate(SW = sw55, SW_lwr = sw55_lwr, SW_upr = sw55_upr, 
                         SW_std = sw55_std, SW_std_lwr = sw55_std_lwr, 
                         SW_std_upr = sw55_std_upr, SW_sd = sw55.sd, SW_std_sd = sw55std.sd) %>% 
                  select(AID, SW, SW_lwr, SW_upr, SW_sd, SW_std, SW_std_lwr, SW_std_upr, SW_std_sd),
                
                
                bc_dat %>% 
                  dplyr::filter(Species == "AMW") %>% 
                  mutate(SW = sw40, SW_lwr = sw40_lwr, SW_upr = sw40_upr, 
                         SW_std = sw40_std, SW_std_lwr = sw40_std_lwr, 
                         SW_std_upr = sw40_std_upr, SW_sd  = sw40.sd, SW_std_sd = sw40std.sd) %>% 
                  select(AID, SW, SW_lwr, SW_upr, SW_sd, SW_std, SW_std_lwr, SW_std_upr, SW_std_sd)
      )

bc_dat <- bc_dat %>% left_join(std_SW, by = "AID")
names(bc_dat)[names(bc_dat) == "SW.x"] <- "SingleWidth"
names(bc_dat)[names(bc_dat) == "SW.y"] <- "SW"

# confirm proper grouping
table(bc_dat$group)
```



## Analysis {.tabset}

1. Compare Body Shapes  
2. How does error scale from 1D to 2D to 3D?  
3. Compare Unstd metrics  
4. Compare std metrics  
5. How are they correlated?  
6. Which is most precise to itself  
7. Altimeter differences?  


### Body Shapes
Figure 2  
Compare Body shapes betweeen blues, AMWs, humpbacks, and humpback calves
```{r}
# Arrange table
rename.widths <-  c("5%","10%", "15%", "20%", "25%", "30%", "35%", "40%", "45%", "50%", "55%","60%","65%", "70%", "75%", "80%", "85%", "90%")

names(bc_dat)[5:22] <- rename.widths


sp.widths <- bc_dat %>% select(names(bc_dat)[5:22],'AID','TL','Image', 'group') %>% 
  pivot_longer(., cols = names(bc_dat)[5:22], names_to = "Var", values_to = "Val") 

sp.widths$WidRatio <- sp.widths$Val/sp.widths$TL
sp.widths$group <- as.factor(sp.widths$group)

#
# Create Body Shape Plot 
#

# Absolute widths (m)
widths_abs <- ggplot(aes(x=Var, y = Val, color = group), data=sp.widths) +
  geom_boxplot() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  theme(legend.position = "none") + 
  scale_x_discrete(name ="width % of TL", limits=rename.widths) +
  facet_wrap(~group) +
  ylab("Width  (m)") + scale_colour_manual(values = specColors)

# Relative widths (standardized Width Ratio)
widths_rel <- ggplot(aes(x=Var, y = WidRatio, color = group), data=sp.widths) +
  geom_boxplot() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) + 
  theme(legend.position = "none") + 
  scale_x_discrete(name ="width % of TL", limits=rename.widths) +
  facet_wrap(~group) +
  ylab("Standardized Width") + scale_colour_manual(values = specColors)

ggarrange(widths_abs,widths_rel, ncol = 1, labels = 'AUTO')
ggsave("figures/Fig2_body_shapes.jpg", width = 180, units = "mm")    
ggsave("figures/Fig2_body_shapes.tiff", width = 180, units = "mm")
```


### Ridgleplot of posterior widths
Used for Figure 1B
```{r}
# pull example AID
id <- indv_df$`BW180829-30`
names(id)[3:20] <- c("5%", "10%", "15%","20%", "25%", "30%", "35%", "40%", "45%", "50%", 
                     "55%", "60%", "65%", "70%", "75%", "80%", "85%", "90%")

id <- id %>% select(names(id)[3:20]) %>% 
  pivot_longer(., cols = names(id)[3:20], names_to = "width_name", values_to = "width") 
id$width_name <- ordered(as.factor(id$width_name), c("90%", "85%","80%","75%","70%","65%","60%","55%","50%","45%",
                                                     "40%","35%","30%","25%","20%","15%", "10%","5%"))


ggplot(data = id, aes(x = width, y = width_name)) + 
  geom_density_ridges(color = "steelblue", fill = "steelblue", alpha = .75, scale = 1.2, rel_min_height = 0.05) + 
  xlab("width (m)") + ylab("") + theme(axis.text=element_text(size=13), 
        axis.title=element_text(size=14))
ggsave("figures/Fig1B_ridgeplot_widths.jpg",  width = 5, height = 9)
ggsave("figures/Fig1B_ridgeplot_widths.tiff",  width = 5, height = 9)
```


### Posterior Distribution Examples
Figure 3  
Show example of an individual's posterior predictive distributions for TL, SW, SWstd, SA, BAI, and BV
```{r}
# select individual blue whale to show example of posterior distributions
ex_output <- indv_df$`BW180829-30`


exTL <- ggplot(ex_output, aes((TL))) + 
  stat_halfeye(fill = "steelblue", alpha = 0.75) + 
  xlab("predicted TL (m)") + 
  ylab("density") + 
    ggtitle("Total length") + 
  theme(plot.title = element_text(size = 10))

exSW <- ggplot(ex_output, aes(TL.55.00..Width)) + 
  stat_halfeye(fill = "steelblue", alpha = 0.75) + 
  xlab("predicted SW (m)") + 
  ylab("density") + 
    ggtitle("1D: SW") + 
  theme(plot.title = element_text(size = 10))

exSWstd <- ggplot(ex_output, aes(SWstd)) + # note that "SW" here was calculated in CollatriX and is already standardized 
  stat_halfeye(fill = "steelblue", alpha = 0.75) + 
  xlab("predicted SWstd") + 
  ylab("density") + 
    ggtitle("1D: SWstd") + 
  theme(plot.title = element_text(size = 10))

exSA <- ggplot(ex_output, aes(SA_HT)) + 
  stat_halfeye(fill = "steelblue", alpha = 0.75) + 
  xlab("predicted SA (m2)") + 
  ylab("density") + 
    ggtitle("2D: Surface Area") + 
  theme(plot.title = element_text(size = 10))

exBAI <- ggplot(ex_output, aes(BAIpar_HT)) + 
  stat_halfeye(fill = "steelblue", alpha = 0.75) + 
  xlab("predicted BAI") +
  ylab("density") + 
    ggtitle("2D: BAI") + 
  theme(plot.title = element_text(size = 10))

exBV <- ggplot(ex_output, aes(BV_HT)) + 
  stat_halfeye(fill = "steelblue", alpha = 0.75) + 
  xlab("predicted BV (m3)") + 
  ylab("density") + 
    ggtitle("3D: Body Volume") + 
  theme(plot.title = element_text(size = 10))


ggarrange(exTL, exSW, exSWstd, exSA, exBAI, exBV)
ggsave("figures/Fig3_posteriors.jpg", width = 180, units = "mm")  
ggsave("figures/Fig3_posteriors.tiff", width = 180, units = "mm")  

```


### Body condition estimates in 1d, 2d, and 3d.
Figure 4  
SW, SA, BV, SWstd, BAI
```{r}
# SW
sw_plot <- ggplot(data = bc_dat, aes(x=TL.mean, y = SW, ymin = SW_lwr, ymax = SW_upr, color = group)) +
  geom_pointrange(fill = "grey70") +
  theme(legend.position = "none") + 
  ylab("Single Width (m)") + xlab("TL (m)") + 
  scale_colour_manual(values = specColors) +
  ggtitle("1D: SW") 

# FDSA using HT
HT_fdsa <- ggplot(data = bc_dat, aes(x=TL.mean, y = SA.HT.mean , ymin = SA.HT.lower, ymax = SA.HT.upper, color = group)) +
  geom_pointrange(fill = "grey70") +
  theme(legend.position = "none") + 
  ylab("SA (m2)") + 
  xlab("TL (m)") + 
  scale_colour_manual(values = specColors) +
  ggtitle("2D: SA") 

# BV using HT
HT_bv1 <- ggplot(data = bc_dat, aes(x=TL.mean, y = BV.HT.mean, ymin = BV.HT.lower, ymax = BV.HT.upper, color = group)) +
  geom_pointrange(fill = "grey70") +
  theme(legend.position = "none") + 
  ylab(" BV (m3)") + 
  #scale_y_continuous(limits = c(-.5, .5)) +
  scale_colour_manual(values = specColors) +
  xlab("TL (m)") + 
  ggtitle("3D: BV") 

# BAI 
bai <- ggplot(data = bc_dat, aes(x=TL.mean, y = BAIpar.HT.mean, ymin = BAIpar.HT.lower, ymax = BAIpar.HT.upper, color = group)) +
  geom_pointrange(fill = "grey70") +
  theme(legend.position = "none") + 
  ylab("BAI") + 
  scale_colour_manual(values = specColors) +
  xlab("TL (m)") +
  ggtitle("2D: BAI") 


std_SW2 <- ggplot(data = bc_dat, 
                  aes(x=TL.mean, y = SW_std, ymin = SW_std_lwr, ymax = SW_std_upr, color = group)) +
  geom_pointrange(fill = "grey70") +
  #theme(legend.position = "none") + 
  ylab("SW standardized") + xlab("TL (m)") + 
  ylim(0, 0.4) +
  scale_colour_manual(values = specColors) +
  ggtitle("1D: SWstd") + 
  theme(legend.position = "none")


# Legend
leg_std <- ggplot(data = bc_dat, aes(x = TL.mean, y = SW, color = group)) +
  geom_point() +
  lims(x = c(0,0), y = c(0,0))+
  theme_void()+
  theme(legend.position = c(0.4,0.5),
        legend.key.size = unit(0.5, "cm"),
        legend.text = element_text(size =  8),
        legend.title = element_blank()) + #, face = "regular"))+
    scale_colour_manual(values = specColors) +
  guides(colour = guide_legend(override.aes = list(size=4.5)))


ggarrange(sw_plot, HT_fdsa, HT_bv1, std_SW2, bai, leg_std) 
ggsave("figures/Fig4_BC_estimates.jpg", width = 10, height = 7)
ggsave("figures/Fig4_BC_estimates.tiff", width = 10, height = 7)
```


### How does uncerainty scale across 1D, 2D, and 3D
Figure 5    
Use the standard deviation of each body condition metric's posterior predictive distribution
```{r}
# create condensed dataframe
bc_scale <- bc_dat %>% select(AID, Species, group, SW_sd, SA.sd, BV.sd, TL.sd)

# SA ~ SW
mod.scale.sa <- lm(data = bc_scale, log(SA.sd) ~ log(SW_sd))
summary(mod.scale.sa)           # slope = 1.45
round(confint(mod.scale.sa),2)  # CI: [1.20   1.69]


# BV ~ SW
mod.scale.bv <- lm(data = bc_scale, log(BV.sd) ~ log(SW_sd))
summary(mod.scale.bv)           # slope = 1.76 ---slope = 2.14
mod.scale.bv$coefficients
round(confint(mod.scale.bv),2)  #CI:  [1.39   2.13]


s <- ggplot(data = bc_scale, aes(x = log(SW_sd), y = log(SA.sd), color = group)) +
  geom_point() + ylab("log(SA uncertainty)") + xlab("log (SW uncertainty)") + 
  #geom_smooth(method = 'lm') + 
  ylim(-2, 2) + 
  ggtitle("1D vs. 2D") + scale_colour_manual(values = specColors) + 
   theme(legend.position = c(0.2,0.85),
        legend.key.size = unit(0.5, "cm"),
        legend.text = element_text(size =  8),
        legend.title = element_blank()) +
  #geom_smooth(method = 'lm') + 
  geom_abline(intercept = mod.scale.sa$coefficients[1], slope = mod.scale.sa$coefficients[2])

b <- ggplot(data = bc_scale, aes(x = log(SW_sd), y = log(BV.sd), color = group)) +
  geom_point() + ylab("log(BV uncertainty)") + xlab("log (SW uncertainty)") + 
  ylim(-2, 2) + 
  ggtitle("1D vs. 3D") + scale_colour_manual(values = specColors) + theme(legend.position = "none") + 
  #geom_smooth(method = 'lm') + 
  geom_abline(intercept = mod.scale.bv$coefficients[1], slope = mod.scale.bv$coefficients[2])

ggarrange(s, b)
ggsave("figures/Fig5_scaling_of_uncertainty.jpg", width = 180, units = "mm")
ggsave("figures/Fig5_scaling_of_uncertainty.tiff", width = 180, units = "mm")
```

Uncertainty in SA and BV both increase proportionally larger than uncertainty associated with single measurements, though BV increases at a greater rate than SA.



### Calculate body condition index (BCI) for SA and BV
```{r}

#
## model BCI:  HT FDSA uncertainty
#

# HT.mean 
mod.bci.HT.SA <- lm(log(SA.HT.mean) ~ log(TL.mean)*group, data = bc_dat)
summary(mod.bci.HT.SA)
bc_dat$res_SA_HT <- mod.bci.HT.SA$residuals 

# lower bound
mod.bci.HT.SA.lwr <- lm(log(SA.HT.lower) ~ log(TL.mean)*group, data = bc_dat)
summary(mod.bci.HT.SA.lwr)
bc_dat$res_SA_HT_lwr <- mod.bci.HT.SA.lwr$residuals 

# upper bound
mod.bci.HT.SA.upr <- lm(log(SA.HT.upper) ~ log(TL.mean)*group, data = bc_dat)
summary(mod.bci.HT.SA.upr)
bc_dat$res_SA_HT_upr <- mod.bci.HT.SA.upr$residuals 


#
## model BCI:  HT BV uncertainty
#

# model BCI HT Uncertainty
mod.bci.HT.BV <- lm(log(BV.HT.mean) ~ log(TL.mean)*group, data = bc_dat)
summary(mod.bci.HT.BV)
bc_dat$res_BV_HT <- mod.bci.HT.BV$residuals 

# lower bound
mod.bci.HT.BV.lwr <- lm(log(BV.HT.lower) ~ log(TL.mean)*group, data = bc_dat)
summary(mod.bci.HT.BV.lwr)
bc_dat$res_BV_HT_lwr <- mod.bci.HT.BV.lwr$residuals 

# upper bound
mod.bci.HT.BV.upr <- lm(log(BV.HT.upper) ~ log(TL.mean)*group, data = bc_dat)
summary(mod.bci.HT.BV.upr)
bc_dat$res_BV_HT_upr <- mod.bci.HT.BV.upr$residuals 
```



#### Summary
for Table 2.
```{r}
metric_tab <- bc_dat %>% group_by(group) %>% summarise(TL.pop.mean = mean(TL.mean),
                                                       TL.pop.sd = sd(TL.mean),
                                                       #TL.range = mean(TL.upper - TL.lower),
                                                       TL.lwr.mean = min(TL.mean),
                                                       TL.upr.mean = max(TL.mean),
                                                       
                                                       SW.mean = mean(SW),
                                         SW.pop.sd = sd(SW),
                                        # SW.range = mean(SW_upr - SW_lwr),
                                         SW.lwr.mean = min(SW_lwr),
                                         SW.upr.mean = max(SW_upr),
                                         
                                         
                                         SWstd.mean = mean(SW_std),
                                         SWstd.pop.sd = sd(SW_std),
                                         #SWstd.range = mean(SW_std_upr - SW_std_lwr),
                                         SWstd.lwr.mean = min(SW_std),
                                         SWstd.upr.mean = mean(SW_std),
                                         
                                         SA.mean = mean(SA.HT.mean),
                                         SA.pop.sd = sd(SA.HT.mean),
                                         #SA.range = mean(SA.HT.upper - SA.HT.lower),
                                         SA.lwr.mean = min(SA.HT.mean),
                                         SA.upr.mean = max(SA.HT.mean),
                                         
                                         BV.mean = mean(BV.HT.mean),
                                         BV.pop.sd = sd(BV.HT.mean),
                                         #BV.range = mean(BV.HT.upper - BV.HT.lower),
                                         BV.lwr.mean = min(BV.HT.mean),
                                         BV.upr.mean = max(BV.HT.mean),
                                         
                                         BAI.mean = mean(BAIpar.HT.mean),
                                         BAI.pop.sd = sd(BAIpar.HT.mean),
                                         #BAI.range = mean(BAIpar.HT.upper - BAIpar.HT.lower),
                                         BAI.lwr.mean = min(BAIpar.HT.mean),
                                         BAI.upr.mean = max(BAIpar.HT.mean)) 


# For Table 2.
metric_tab
```


### Within-metric Precision
Figure 6  
Compare precision across body condition measurements using coefficient of variation (CV%).
The CV% statistic is a simple and widely-used standardized measure of the spread of a set of measurements of a sample. It is a useful value because it allows direct comparison of the variation in samples measured with different units, or with very different means. The CV% is sometimes preferred to the standard deviation because the value of the coefficient of variation is independent of the unit of measurement scale (as long as it is a ratio scale). 
```{r}

# CV% for each metric
cv_dat <- bc_dat %>% mutate(SW = (SW_sd/SW)*100,
                            SWstd = (SW_std_sd/SW_std)*100, 
                            BAI = (BAI.sd/BAIpar.HT.mean)*100,
                            SA = (SA.sd/SA.HT.mean)*100,
                            BV = (BV.sd/BV.HT.mean)*100) %>% select(AID, Species, SW, SWstd, BAI, SA, BV)

cv_piv <- cv_dat  %>% select(names(cv_dat)[3:7],'AID','Species') %>% pivot_longer(., cols = names(cv_dat)[3:7], names_to = "Method", values_to = "cv") 

#cv_piv$Method <- factor(cv_piv$Method, ordered = FALSE)
cv_piv$Method <- ordered(cv_piv$Method, levels = c("SW", "SWstd", "BAI", "SA", "BV"))

ggplot(data = cv_piv, aes(x = Method, y = cv)) + geom_boxplot() + ylab("CV %") + xlab("")    
ggsave("figures/Fig6_CVs.jpg", width = 85, units = "mm")
ggsave("figures/Fig6_CVs.tiff", width = 85, units = "mm")

# CV% summary table (Table 3)
cv_piv %>% group_by(Method) %>% summarize(mean = round(mean(cv),2),
                                          min = round(min(cv),2),
                                          max = round(max(cv),2),
                                          sd = round(sd(cv), 2))

# CV% by species 
ggplot(data = cv_piv, aes(x = Method, y = cv, color = Species)) + geom_boxplot() + ylab("CV %") + xlab("") 
```




### Correlation matrix
for Figure 7  
Comparison of standardized body condition measurements  

#### A) Correlelogram
```{r}
# Helpful links
    # http://www.sthda.com/english/wiki/correlation-matrix-a-quick-start-guide-to-analyze-format-and-visualize-a-correlation-matrix-using-r-software
    # http://www.sthda.com/english/wiki/visualize-correlation-matrix-using-correlogram


# Pearson correlation (r), which measures a linear dependence between two variables (x and y). 
# covariance measures the direction of the linear relationship between two variables.
# correlation measures the strength and direction of the linear relationship

#
# standardized body condition measurement
#
# blue whales
dat.blue <- bc_dat %>% dplyr::filter(Species == "Blue") %>% select(c(SW_std, BAIpar.HT.mean, res_SA_HT, res_BV_HT))
names(dat.blue) <- c("SWstd", "BAI", "SA_BCI", "BV_BCI")
res.blue <- cor(dat.blue)
round(res.blue,2)

# humpback whales
  # adults
dat.hump <- bc_dat %>% dplyr::filter(Species == "Humpback" & Reproductive_Class != "calf") %>% select(c(SW_std, BAIpar.HT.mean, res_SA_HT, res_BV_HT))
names(dat.hump) <- c("SWstd", "BAI", "SA_BCI", "BV_BCI")
res.hump <- cor(dat.hump)
round(res.hump,2)
  # calves
dat.hump.calf <- bc_dat %>% dplyr::filter(Species == "Humpback" & Reproductive_Class == "calf") %>% select(c(SW_std, BAIpar.HT.mean, res_SA_HT, res_BV_HT))
names(dat.hump.calf) <- c("SWstd", "BAI", "SA_BCI", "BV_BCI")
res.hump.calf <- cor(dat.hump.calf)
round(res.hump.calf,2)

# minke whales
dat.minke <- bc_dat %>% dplyr::filter(Species == "AMW") %>% select(c(SW_std, BAIpar.HT.mean, res_SA_HT, res_BV_HT))
names(dat.minke) <- c("SWstd", "BAI", "SA_BCI", "BV_BCI")
res.minke <- cor(dat.minke)
round(res.minke,2)

#
# Correlation matrix (correlogram) w/ p-value
#
#  blue colors = positively correlated, red colors = negatively correlated

pdf(file="figures/Fig7A_correlation_matrix.pdf")
par(mfrow=c(2,2))

corrplot(res.minke, method="number", type = "upper", #order = "hclust", 
         tl.col = "black", tl.srt = 45, diag=FALSE, title = "Antarctic Minkes", mar=c(0,0,1,0))

corrplot(res.blue, method="number", type = "upper", #order = "hclust", 
         tl.col = "black", tl.srt = 45, diag=FALSE, title = "Blues", mar=c(0,0,1,0))

corrplot(res.hump, method="number", type = "upper", #order = "hclust", 
         tl.col = "black", tl.srt = 45, diag=FALSE, title = "Humpbacks", mar=c(0,0,1,0))

corrplot(res.hump.calf, method="number", type = "upper", #order = "hclust", 
         tl.col = "black", tl.srt = 45, diag=FALSE, title = "Humpback calves", mar=c(0,0,1,0))
dev.off()
```


#### B) Regression 
```{r}
cor1 <- ggplot(data = bc_dat, aes(x = BAIpar.HT.mean, y = SW_std, color = group)) + geom_point() + 
  geom_smooth(method = 'lm') + scale_colour_manual(values = specColors) +
  xlab("BAI") + ylab("standardized SW") + theme(legend.position = "none")

cor2 <- ggplot(data = bc_dat, aes(x = BAIpar.HT.mean, y = res_SA_HT, color = group)) + geom_point() + 
  geom_smooth(method = 'lm')  + scale_colour_manual(values = specColors) +
  xlab("BAI") + ylab("BCI: SA") + theme(legend.position = "none")

co3 <- ggplot(data = bc_dat, aes(x = BAIpar.HT.mean, y = res_BV_HT, color = group)) + geom_point() + 
  geom_smooth(method = 'lm') +   scale_colour_manual(values = specColors) +
  xlab("BAI") + ylab("BCI: BV") + theme(legend.position = "none")

ggarrange(cor1,cor2,co3, leg_std)
ggsave("figures/Fig7B_correlations.jpg", width = 7, height = 7) 
```



